---
title: "R masterscript for humann2 data"
output: html_notebook
---

Loading library
```{r}
library(tidyverse)
library(cowplot) # required to arrange multiple plots in a grid
library(RColorBrewer)
library(ggExtra)
theme_set(theme_bw(base_size=12)) # set default ggplot2 theme
```

Loading data

Loading pathway abundance data
```{r}
path_abundance <- read_delim("./pathabundance.tsv", 
    "\t", escape_double = FALSE, trim_ws = TRUE)
glimpse(path_abundance)
```




Moving file id from row to column (pivoting)
```{r}
pivot_path_abundance <- path_abundance %>% pivot_longer(-Pathway, names_to = "sample_id", values_to = "Abundance")
```

Deleting parts of sample id name
```{r}
pivot_path_abundance = pivot_path_abundance %>%
  mutate(sample_id = gsub("merged_([A-Z]+[0-9]+)_Abundance","\\1", pivot_path_abundance$sample_id))
  
head(pivot_path_abundance)
```

# Delete samples SRR4838360, SRR4838361, SRR5479799

```{r}

pivot_path_abundance <- pivot_path_abundance [!(pivot_path_abundance$sample_id=="SRR4838360" | pivot_path_abundance$sample_id=="SRR4838361" | pivot_path_abundance$sample_id=="SRR5479799" ),]

head(pivot_path_abundance)

```

Abundance norm, adding extra column
```{r}
 mutated_path_abundance <- pivot_path_abundance %>%
  group_by(sample_id) %>%
  mutate(abundance_norm = Abundance / sum(Abundance)) %>%
  ungroup()

head(mutated_path_abundance)

mutated_path_abundance %>%
  group_by(sample_id) %>%
  summarize(sum_abundance_norm = sum(abundance_norm)) %>%
  ungroup() %>%
  head()

```
Log abundance_norm
```{r}
path_abundance_noBact = mutated_path_abundance %>% 
  mutate(log_abundance_norm = log10(abundance_norm + 1e-10)) %>% 
  filter(Pathway != "UNMAPPED") %>% 
  filter(!grepl("UNINTEGRATED",Pathway)) %>%
  filter(!grepl("\\|",Pathway))
head(path_abundance_noBact)
  
```
Plotting number of zero pathways against sample
```{r}
p_barplot_path = path_abundance_noBact %>% 
  group_by(sample_id) %>% 
  filter(Abundance != 0) %>%
  summarise(number_nonZero_pathways = length(Pathway)) %>% 
  ungroup() %>%
  ggplot(aes(x=sample_id, y = number_nonZero_pathways)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  ylab("Number of non-zero pathways") +
  xlab("Sample") +
  theme(axis.text.x = element_text(angle = 90))

p_barplot_path

```





# Global analysis
Join all meta files(1-5)
```{r}
meta <- bind_rows(meta1, meta2, meta3, meta4_2, meta5_2)
meta

#exported this meta file, deleted three amplicon samples and imported again below.


meta_3deleted <- read_delim("./meta_3deleted_samples.txt", 
    "\t", escape_double = FALSE, trim_ws = TRUE)
glimpse(meta_3deleted)
```
Adding id column e.g. s01_soi_gup (number_environment_study)
```{r}
meta_3deleted <- meta_3deleted %>% 
  mutate(id = paste0("s",sprintf("%02d", 1:nrow(meta_3deleted)))) %>% 
  mutate(id = case_when(
    environment == "soil metagenome" ~ paste0(id,"_soilm_"),
    environment == "wastewater metagenome" ~ paste0(id,"_waste_"),
    environment == "marine metagenome" ~ paste0(id,"_marin_"),
    environment == "freshwater sediment metagenome" ~ paste0(id,"_frwas_"),
    environment == "freshwater metagenome" ~ paste0(id,"_fresh_"),
    
  ))


meta_3deleted <- meta_3deleted %>% 
  mutate(id = case_when(
    study == "J. Gupta et al. (2017)" ~ paste0(id, "gup"),
    study == "E. Ransom-Jones et al. (2017)" ~ paste0(id, "ran"),
    study == "L.J. Pinnell, J.W. Turner (2019)" ~ paste0(id, "pin"),
    study == "M.L. Petrovich et al. (2019)" ~ paste0(id, "pet"),
    study == "J.A. Bryant et al. (2016)" ~ paste0(id, "bry")
 ))


meta_3deleted
    
```
Rename environments
```{r}
meta_3deleted <- meta_3deleted %>% 
mutate(environment = case_when(
    environment == "soil metagenome" ~ "soil landfill, India",
    environment == "wastewater metagenome" ~ "wastewater landfill, UK",
    environment == "marine metagenome" ~ "marine, Gulf of Mexico",
    environment == "freshwater sediment metagenome" ~ "marine, North Pacific Ocean",
    environment == "freshwater metagenome" ~ "wastewater treatment site, US",
    
  ))

meta_3deleted
```

Changing order of columns in meta
```{r}
meta_3deleted <- meta_3deleted %>% 

select(id, sample_id, sample_alias, environment, study)

meta_3deleted
```


Join path abundance with meta
```{r}

join_path_meta <- left_join(path_abundance_noBact, meta_3deleted, by='sample_id') 
              
head(join_path_meta)

```
Changing order of columns in merged file (join_path_meta)
```{r}
join_path_meta <- join_path_meta %>% 

select(id, sample_id, sample_alias, environment, Pathway, study, Abundance, abundance_norm, log_abundance_norm)

head(join_path_meta)
```




#PCA
Pivot wider (id, Pathway, abundance_norm)
```{r}
select_path_meta <- join_path_meta %>% 
select(id, Pathway, log_abundance_norm)
head(select_path_meta)

pivot_path_meta <- select_path_meta %>% 
pivot_wider(names_from = id, values_from = log_abundance_norm)

head(pivot_path_meta)



```
Changing pathway from text to numerical values
```{r}

df_pathway <- pivot_path_meta[,-1]
df_pathway = as.data.frame(df_pathway, row.names = NULL, optional = FALSE,)
row.names(df_pathway) <- pivot_path_meta$Pathway
df_pathway = t(df_pathway)

glimpse(df_pathway)
```



Preparing PCA components
```{r}

pca <- prcomp(df_pathway) # center = TRUE, scale. = TRUE)

summary(pca)
  
```

Add study + environment information back into PCA data
```{r}

pca_data = cbind(id = rownames(pca$x[,1:2]),
                 pca$x[,1:2]) %>%
  as_tibble() %>% 
  left_join(meta_3deleted, by = "id")  %>% 
  mutate(PC1 = as.double(PC1), 
         PC2 = as.double(PC2))

glimpse(pca_data)


```

Adding percentage of variance captured
```{r}
percentage <- round(pca$sdev / sum(pca$sdev) * 100, 2)
percentage <-
  paste0(colnames(pca$x), " (", paste(as.character(percentage), "%", ")"))
head(percentage)

percentage[1]
```

Plot PC1 & PC2, colored by study
```{r}
pca_data %>% 
ggplot(aes(PC1, PC2, group = study, color = study)) +
  geom_point() +
  stat_ellipse() +
  xlab(percentage[1]) +
  ylab(percentage[2])
```
Plot PC1 & PC2, colored by environment
```{r, fig.width= 10, fig.height= 4}
p_pca_env = pca_data %>% 
  ggplot(aes(PC1, PC2, group = environment, color = environment)) +
  geom_point(size = 1.5) +
  stat_ellipse(size = 1, alpha = 0.2, geom = "polygon", aes(fill = environment)) +
  xlab(percentage[1]) +
  ylab(percentage[2]) +
  theme_classic() +
  scale_color_brewer(type = "qual", palette = "Set1") +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  theme(legend.position = "left",
        aspect.ratio = 1)

p_pca_env = ggMarginal(p_pca_env, groupColour = TRUE)

p_pca_env

ggsave(filename = "PCA", plot = p_pca_env, device = "pdf")

```

# Find missing samples: SRR4838360, SRR4838361, SRR5479799 "amplicon" (From E. Ransom Jones)

```{r}
pivot_path_abundance %>% 
  inner_join(meta_3deleted) %>% 
  select(sample_id) %>% 
  unique()

samples_meta = meta_3deleted$sample_id
samples_path = pivot_path_abundance %>% 
  select(sample_id) %>% 
  unique() %>% 
  .$sample_id

missing_samples = setdiff(samples_meta,samples_path)

meta_3deleted %>% 
  filter(sample_id %in% missing_samples)

```
#PCA - Select rows that include pathways that degrade plastic
```{r}
plastic_rows = pivot_path_meta %>%
  slice(110, 308, 357)

plastic_rows
```
Changing pathway from text to numerical values
```{r}

df_plastic_rows <- plastic_rows[,-1]
df_plastic_rows = as.data.frame(df_plastic_rows, row.names = NULL, optional = FALSE,)
row.names(df_plastic_rows) <- plastic_rows$Pathway
df_plastic_rows = t(df_plastic_rows)

glimpse(df_plastic_rows)
```



Preparing PCA components for plastic pathways
```{r}

pca_plastic <- prcomp(df_plastic_rows) # center = TRUE, scale. = TRUE)

summary(pca_plastic)
  
```

Add study + environment information back into PCA plastic pathways data
```{r}

pca_plastic_data = cbind(id = rownames(pca_plastic$x[,1:2]),
                 pca_plastic$x[,1:2]) %>%
  as_tibble() %>% 
  left_join(meta_3deleted, by = "id")  %>% 
  mutate(PC1 = as.double(PC1), 
         PC2 = as.double(PC2))

glimpse(pca_plastic_data)

```
Adding percentage of variance captured
```{r}
percentage_plastic <- round(pca_plastic$sdev / sum(pca_plastic$sdev) * 100, 2)
percentage_plastic <-
  paste0(colnames(pca_plastic$x), " (", paste(as.character(percentage_plastic), "%", ")"))
head(percentage_plastic)

percentage_plastic[1]
```

Plot PC1 & PC2, colored by environment for plastic pathways
```{r, fig.width= 10, fig.height= 4}
p_pca_env_plastic = pca_plastic_data %>% 
  ggplot(aes(PC1, PC2, group = environment, color = environment)) +
  geom_point(size = 1.5) +
  stat_ellipse(size = 1, alpha = 0.2, geom = "polygon", aes(fill = environment)) +
  xlab(percentage_plastic[1]) +
  ylab(percentage_plastic[2]) +
  theme_classic() +
  scale_color_brewer(type = "qual", palette = "Set1") +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  theme(legend.position = "bottom",
        aspect.ratio = 1)

p_pca_env_plastic = ggMarginal(p_pca_env_plastic, groupColour = TRUE)

p_pca_env_plastic

```
#Violin plot for plastic pathways
"P184-PWY: protocatechuate degradation I (meta-cleavage pathway)" 
"PWY-6435: 4-hydroxybenzoate biosynthesis V" 
"PWY-7002: 4-hydroxyacetophenone degradation"
```{r}
plastic_PWY_6435 = join_path_meta %>% 
 filter(grepl("PWY-6435", Pathway, fixed = TRUE))
plastic_PWY_6435
```


"PWY-6435: 4-hydroxybenzoate biosynthesis V" 
```{r}
p_violin_plastic_PWY_6435 = plastic_PWY_6435%>% 
ggplot(aes(x=environment, y=log_abundance_norm, fill=environment)) +
  geom_violin(trim=FALSE) +
  geom_jitter(alpha = 0.3) +
  theme_classic() +
  scale_color_brewer(type = "qual", palette = "Set1") +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  ggtitle('PWY-6435: 4-hydroxybenzoate biosynthesis V') +
  ylab("Log abundance norm") +
  xlab("Environment") +
  theme(axis.text.x = element_text(angle = 90)) 
p_violin_plastic_PWY_6435
ggsave(filename = "PWY_6435.pdf", plot = p_violin_plastic_PWY_6435, device = "pdf")
```

"P184-PWY: protocatechuate degradation I (meta-cleavage pathway)" 
```{r}
plastic_P184_PWY = join_path_meta %>% 
 filter(grepl("P184-PWY", Pathway, fixed = TRUE))
plastic_P184_PWY
```
"P184-PWY: protocatechuate degradation I (meta-cleavage pathway)" 
```{r}
p_violin_plastic_P184_PWY = plastic_P184_PWY%>% 
ggplot(aes(x=environment, y=log_abundance_norm, fill=environment)) +
  geom_violin(trim=FALSE) +
  geom_jitter(alpha = 0.3) +
  theme_classic() +
  scale_color_brewer(type = "qual", palette = "Set1") +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  ggtitle('P184-PWY: protocatechuate degradation I (meta-cleavage pathway)') +
  ylab("Log abundance norm") +
  xlab("Environment") +
  theme(axis.text.x = element_text(angle = 90)) 
p_violin_plastic_P184_PWY
ggsave(filename = "PWY_P184.pdf", plot = p_violin_plastic_P184_PWY, device = "pdf")
```

"PWY-7002: 4-hydroxyacetophenone degradation"
```{r}
plastic_PWY_7002 = join_path_meta %>% 
 filter(grepl("PWY-7002", Pathway, fixed = TRUE))
plastic_PWY_7002
```
"PWY-7002: 4-hydroxyacetophenone degradation"
```{r}
p_violin_plastic_PWY_7002 = plastic_PWY_7002%>% 
ggplot(aes(x=environment, y=log_abundance_norm, fill=environment)) +
  geom_violin(trim=FALSE) +
  geom_jitter(alpha = 0.3) +
  theme_classic() +
  scale_color_brewer(type = "qual", palette = "Set1") +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  ggtitle('PWY-7002: 4-hydroxyacetophenone degradation') +
  ylab("Log abundance norm") +
  xlab("Environment") +
  theme(axis.text.x = element_text(angle = 90)) 
p_violin_plastic_PWY_7002

ggsave(filename = "PWY_7002.pdf", plot = p_violin_plastic_PWY_7002, device = "pdf")

```