---
title: "Visualize fluxes from random sampling"
author: "iGEM 2020"
output:
  html_document:
    df_print: kable
    toc: true
    toc_float: true
---
 
# Load libraries
```{r}
library(tidyverse)
library(cowplot)
library(ggpubr)
library(ggsci)
library(ggrepel)
library(ggExtra)
library(readxl)
library(ComplexHeatmap)
library(Rtsne)
library(janitor)
library(hrbrthemes)
library(wesanderson)
library(tidyHeatmap)
select = dplyr::select
mutate = dplyr::mutate
```

# Load Data
Load pre-processed vcf data and metadata
```{r}
df_glu <- read_excel("/Users/castilln/Desktop/igem/github/GEM/scripts/scrap/randomSolutionsGlu.xlsx")
df_pet <- read_excel("/Users/castilln/Desktop/igem/github/GEM/scripts/scrap/randomSolutionsPET.xlsx")
df_pu <- read_excel("/Users/castilln/Desktop/igem/github/GEM/scripts/scrap/randomSolutionsPU.xlsx")
 
df_kegg = read_csv("/Users/castilln/Desktop/igem/github/GEM/Data/kegg_df.csv")
```
 
```{r}
df_full = 
 full_join(df_glu,df_pu, by = "rxns_names")

df_full = full_join(df_full, df_pet, by="rxns_names")
head(df_full)
```
Let's try only Glu vs PU 
```{r}
#df_full = full_join(df_glu, df_pu, by = "rxns_names")
```


```{r}
df_long =
  df_full %>% 
  pivot_longer(cols = -rxns_names,
               names_to = "solution",
               values_to = "flux") %>% 
  distinct() %>% 
  mutate(flux = ifelse(is.na(flux),0,flux))
```

Select unique flux distributions
```{r}
df_long %>% 
  group_by(rxns_names, flux) %>% 
  arrange(rxns_names, flux, solution) %>% 
  slice(1L) -> filter_sol
```
 
```{r}
filter_sol %>% ungroup() %>%  select(solution) %>% distinct() -> filter_only_sol
```

```{r}
df_long %>% inner_join(filter_only_sol, by = "solution") -> df_long_unique
```
 
 
```{r}
df_long_unique %>% 
  filter(rxns_names == "BIOMASS_Ec_iML1515_core_75p37M")
```
 
 
Get samples with non-zero biomass flux
```{r}
nonZero_samples = 
  df_long_unique %>% 
  filter(rxns_names == "BIOMASS_Ec_iML1515_core_75p37M") %>% 
  filter(flux > 0)
 
head(nonZero_samples)
```
Filter out all non zero samples, add pathway information from KEGG
```{r}
df_long_nonzero = 
  nonZero_samples %>% 
  select(solution) %>% 
  left_join(df_long_unique, by = c("solution" = "solution")) %>% 
  left_join(df_kegg, by = c("rxns_names" = "rxn_id")) %>% 
  group_by(solution) %>% 
  mutate(NA_fluxes = ifelse(sum(is.na(flux)) > 0,TRUE,FALSE)) %>% 
  ungroup() %>% 
  filter(NA_fluxes == FALSE)
```


Convert into wide format
```{r}
df_wide_nonzero =
  df_long_nonzero %>% 
  select(-c(gene_id,pathway_id)) %>% 
  pivot_wider(names_from = solution,
              values_from = flux) %>% 
  as.data.frame() 
 
rownames(df_wide_nonzero) = df_wide_nonzero[,1]
df_wide_nonzero = df_wide_nonzero[,-1]
```
 
### Dimnesionality reduction
PCA
```{r}
pca1 = prcomp(t(df_wide_nonzero),center = TRUE)
 
plotData = pca1$x[,1:2]
plotData = cbind(solution = rownames(plotData),plotData)
rownames(plotData) = NULL
 
head(plotData)
 
sample_names = plotData[,1]
 
plotData %>% 
  as_tibble() %>% 
  mutate(Condition = case_when(
    grepl("Glu",solution) ~ "Glu",
    grepl("PU",solution) ~ "PU",
    grepl("PET",solution) ~ "PET" 
    )) %>% 
  filter(!is.na(Condition)) %>% 
  mutate(PC1 = as.double(PC1),
         PC2 = as.double(PC2))  %>%
  ggplot(aes(x = PC1,y = PC2, color = Condition)) +
  geom_point() +
  scale_y_continuous(breaks = NULL) + 
  scale_x_continuous(breaks = NULL)
```
 
tSNE
```{r}
tsne1 = Rtsne(t(df_wide_nonzero),
             dims = 2,
             perplexity = 30,
             theta = 0.5,
             check_duplicates = FALSE)
 
tsne1$Y
 
plotData =
  tsne1$Y %>% 
  as_tibble() %>% 
  mutate(solution = sample_names)
 
plotData = 
  plotData %>% 
  mutate(Condition = case_when(
    grepl("Glu",solution) ~ "Glu",
    grepl("PET",solution) ~ "PET",
    grepl("PU",solution) ~ "PU",
    )) %>% 
  mutate(V1 = as.double(V1),
         V2 = as.double(V2))
 
plotData %>% 
  ggplot(aes(x = V1,y = V2, color = Condition, group = Condition)) +
  geom_point() +
  xlab("tSNE 1") +
  ylab("tSNE 2") +
  theme_bw() +
  theme(legend.position = "bottom",
        aspect.ratio = 1,
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        plot.title = element_text(face = "bold")) +
  scale_color_manual("Carbon source", values = c("darkorange1","deeppink4", "darkblue"))
```
 
Hierarchical clustering and heatmaps
```{r}
# Extract nonzero fluxes
nonZero_fluxes = 
  df_long_nonzero %>% 
  group_by(rxns_names) %>% 
  summarize(zeroFlux = sum(flux) == 0) %>%
  ungroup() %>% 
  filter(zeroFlux == FALSE)
 
# Format for Heatmap
hm.data = df_long_nonzero %>% 
  right_join(nonZero_fluxes,by = "rxns_names") %>% 
  select(rxns_names,solution,flux) %>% 
  pivot_wider(names_from = solution,
              values_from = flux) %>% 
  arrange(rxns_names) %>% 
  as.data.frame()
 
rownames(hm.data) = hm.data[,1]
hm.data = hm.data[,-1]
 
hm.data = as.matrix(hm.data)
 
# Get sample data for annotation
df_long_nonzero_cond = 
  df_long_nonzero %>% 
  mutate(Condition = case_when(
    grepl("Glu",solution) ~ "Glu",
    grepl("PET",solution) ~ "PET",
    grepl("PU",solution) ~ "PU",
    ))

sample.cond = 
  df_long_nonzero_cond %>% 
  pull(Condition)
 
pathway.annot = df_long_nonzero %>% 
  right_join(nonZero_fluxes,by = "rxns_names") %>% 
  select(rxns_names,pathway_id) %>%
  distinct() %>% 
  arrange(rxns_names) %>% 
  pull(pathway_id)


#z-score normalization
df_long_nonzero_cond %>% 
  group_by(solution,Condition) %>% 
  mutate(z_score_group = (flux - mean(flux)) / sd(flux)) %>% 
  ungroup() -> df_long_nonzero_cond_norm

hm.data = df_long_nonzero_cond_norm %>% 
  right_join(nonZero_fluxes,by = "rxns_names") %>% 
  select(rxns_names,solution,z_score_group) %>% 
  pivot_wider(names_from = solution,
              values_from = z_score_group) %>% 
  arrange(rxns_names) %>% 
  as.matrix()
 
# Create annotations for heatmap
ha.col = HeatmapAnnotation(Carbon_source = sample.cond,
                       col = list(Carbon_source = c(Glucose = "darkorange1", Polyurethane = "violetred", PET = "darkblue")))

my_palette <- colorRampPalette(c("yellow", "orange", "red")) (n=20) 
breaks <- seq(min(hm.data, na.rm = T), max(hm.data, na.rm = T), length.out = 21)

hm = heatmap.2(hm.data, 
               hm.data, 
               main = "Heatmap of fluxes on different C sources",
               trace = "none",
               col = my_palette,
               breaks = breaks, 
               dendogram = "row",
               ColV = "NA")

hm



# Draw heatmap
hm = Heatmap(hm.data,
        jitter = TRUE,
        name = "scaled-centered intesities",
        #clustering_distance_rows = "euclidean", 
        clustering_distance_columns = "euclidean",
        show_column_names = FALSE,
        show_row_names = FALSE,
        #top_annotation = ha.col,
        #left_annotation = ha.row,
        #col = col_fun,
        #column_dend_height = unit(25, "mm"),
        #row_dend_width = unit(25, "mm"),
        show_row_dend = F,
        show_column_dend = TRUE)
 
hm
```

heatmap.2
```{r}

#try to get rid of nan values
dist_no_na <- function(hm.data) {
    edist <- dist(hm.data)
    edist[which(is.na(edist))] <- max(edist, na.rm=TRUE) * 1.1 
    return(edist)
}

hm = heatmap.2(hm.data, 
               hm.data, 
               main = "Heatmap of fluxes on different C sources",
               trace = "none",
               distfun = dist_no_na(),
               col = my_palette,
               breaks = breaks, 
               dendogram = "column",
               Rowv = "NA",
               na.color = "black")

hm

```


Tidy heatmap
```{r}
options(expression = 800000)
df_long_nonzero_cond_norm %>% 
  as_tibble() %>% 
  filter(z_score_group > 0) %>% 
  heatmap(rxns_names, solution, z_score_group, breaks)
```


Wilcoxon tests
```{r, warning=F}
options(scipen = 1)
res_wilcox =
  df_long_nonzero %>%
  right_join(nonZero_fluxes, by = "rxns_names") %>%
  mutate(Condition = ifelse(grepl("Glu", solution), "Glucose", "Polyurethane")) %>%
  select(rxns_names, flux, Condition) %>%
  group_by(Condition, rxns_names) %>%
  summarise(flux = list(flux)) %>%
  pivot_wider(names_from = Condition, values_from = flux) %>%
  group_by(rxns_names) %>%
  mutate(
    p_value = wilcox.test(unlist(Glucose), unlist(Polyurethane))$p.value,
    fold_change = sign(median(unlist(Polyurethane), median(unlist(Glucose)))) * 
      abs(median(unlist(Polyurethane)) - median(unlist(Glucose)))
  ) %>%
  select(rxns_names, p_value, fold_change) %>%
  mutate(FDR = p.adjust(p_value, method = "BH")) %>%
  filter(FDR < 0.05) %>%
  arrange(FDR)
```
 
Show wilcox test results
```{r}
df_long_nonzero %>% 
  right_join(nonZero_fluxes, by = "rxns_names") %>%
  select(rxns_names, pathway_id) %>% 
  inner_join(res_wilcox, by = "rxns_names") %>% 
  distinct() %>% 
  arrange(FDR)
```
Filter out those with pathways
```{r}
df_long_nonzero %>% 
  right_join(nonZero_fluxes, by = "rxns_names") %>%
  select(rxns_names, pathway_id) %>% 
  inner_join(res_wilcox, by = "rxns_names") %>% 
  distinct() %>%
  arrange(fold_change) %>% 
  filter(!is.na(pathway_id))
```
 
```{r}
rxn_to_plot = "PE161abcpp"
 
df_long_nonzero %>%
  right_join(nonZero_fluxes, by = "rxns_names") %>%
  mutate(Condition = case_when(
    grepl("Glu",solution) ~ "Glu",
    grepl("PET",solution) ~ "PET",
    grepl("PU",solution) ~ "PU",
    )) %>%
  filter(rxns_names == rxn_to_plot) %>% 
  ggplot(aes(x=Condition,y=flux,color = Condition, fill = Condition)) +
  geom_jitter() +
  #theme_ipsum_rc(axis = TRUE,
    #             axis_title_face = "bold") +
  ggtitle("PE161abcpp") +
  xlab("") +
  ylab("Flux (mmol/gDWh)") +
  scale_y_continuous(breaks = NULL) +
  theme(legend.position = "none") +
  scale_color_manual(values = wes_palette("Cavalcanti1", n = 3, type = "discrete")) +
  scale_fill_manual(values = wes_palette("Cavalcanti1", n = 3, type = "discrete"))
```


stats kw 2 sided. 
```{r, fig.width=6,fig.height=6}
#pathway_to_plot = "Oxidative phosphorylation" 
#pathway_to_plot = "Glyoxylate and dicarboxylate metabolism" 
#pathway_to_plot = "Glycine, serine and threonine metabolism"
#pathway_to_plot = "Cysteine and methionine metabolism"
#pathway_to_plot = "Pentose phosphate pathway"
#pathway_to_plot = "Pyruvate metabolism"
#pathway_to_plot = "Phenylalanine, tyrosine and tryptophan biosynthesis"
#pathway_to_plot = "Glycolysis / Gluconeogenesis"
pathway_to_plot = "Citrate cycle (TCA cycle)"


df_long_nonzero %>%
  right_join(nonZero_fluxes, by = "rxns_names") %>%
  mutate(Condition = case_when(
    grepl("Glu",solution) ~ "Glu",
    grepl("PET",solution) ~ "PET",
    grepl("PU",solution) ~ "PU",
    )) %>%
  filter(pathway_id == pathway_to_plot) %>% 
  ggplot(aes(x=Condition,y=flux,color = Condition, fill = Condition)) +
  geom_jitter() +
  geom_hline(yintercept = 0) +
  facet_wrap(~rxns_names,
             scales = "free") +
  #theme_ipsum_rc(axis = TRUE,
   #             axis_title_face = "bold") +
  ggtitle("TCA Cycle") +
  xlab("") +
  ylab("Flux (mmol/gDWh)") +
  scale_y_continuous(breaks = NULL) +
  theme_bw() +
  theme(legend.position = "bottom",
        aspect.ratio = 0.7,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(), 
        plot.title = element_text(face = "bold")) +
  scale_color_manual(values = wes_palette("Cavalcanti1", n = 3, type = "discrete")) +
  scale_fill_manual(values = wes_palette("Cavalcanti1", n = 3, type = "discrete")) +
  stat_compare_means(label = "p.signif",vjust = 2, hjust = 1) -> p_tca
  
ggsave("p_tca.png", device = "png", width = 30, height = 25, units = "cm", dpi = "retina")

```
```{r}
rxn_to_plot = "FRD3"
 
df_long_nonzero %>%
  right_join(nonZero_fluxes, by = "rxns_names") %>%
  mutate(Condition = case_when(
    grepl("Glu",solution) ~ "Glu",
    grepl("PET",solution) ~ "PET",
    grepl("PU",solution) ~ "PU",
    )) %>%
  filter(rxns_names == rxn_to_plot) %>% 
  ggplot(aes(x=Condition,y=flux,color = Condition, fill = Condition)) +
  geom_jitter() +
  facet_wrap(~rxns_names,
             scales = "free") +
  #theme_ipsum_rc(axis = TRUE,
   #             axis_title_face = "bold") +
  xlab("") +
  ylab("Flux (mmol/gDWh)") +
  theme(legend.position = "none") +
  scale_color_manual(values = wes_palette("Cavalcanti1", n = 3, type = "discrete")) +
  scale_fill_manual(values = wes_palette("Cavalcanti1", n = 3, type = "discrete"))
```

 
```{r, fig.width=10,fig.height=10}
pathway_to_plot = "Pyruvate metabolism"
pathway_to_plot = "Phenylalanine, tyrosine and tryptophan biosynthesis"
rxn_to_plot = "PGK"
 
df_long_nonzero %>%
  right_join(nonZero_fluxes, by = "rxns_names") %>%
  mutate(Condition = case_when(
    grepl("Glu",solution) ~ "Glu",
    grepl("PET",solution) ~ "PET",
    grepl("PU",solution) ~ "PU",
    )) %>%
  filter(pathway_id == pathway_to_plot) %>% 
  ggplot(aes(x=Condition,y=flux,color = Condition, fill = Condition)) +
  geom_jitter() +
  geom_hline(yintercept = 0) +
  facet_wrap(~rxns_names,
             scales = "free") +
  ggtitle("Pyruvate metabolism") +
  xlab("") +
  ylab("Flux (mmol/gDWh)") +
  scale_y_continuous(breaks = NULL) +
  theme_bw() +
  theme(legend.position = "bottom",
        aspect.ratio = 0.7,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank()) +
  scale_color_manual(values = wes_palette("Cavalcanti1", n = 3, type = "discrete")) +
  scale_fill_manual(values = wes_palette("Cavalcanti1", n = 3, type = "discrete")) +
  stat_compare_means(label = "p.signif",vjust = 2, hjust = 1) -> p_pyruvate
  
  ggsave("p_pyruvate.png", device = "png", width = 30, height = 25, units = "cm", dpi = "retina")

```


```{r}
pathway_to_plot = "Porphyrin and chlorophyll metabolism"
 
df_long_nonzero %>%
  right_join(nonZero_fluxes, by = "rxns_names") %>%
  mutate(Condition = case_when(
    grepl("Glu",solution) ~ "Glu",
    grepl("PET",solution) ~ "PET",
    grepl("PU",solution) ~ "PU",
    )) %>%
  filter(pathway_id == pathway_to_plot) %>% 
  ggplot(aes(x=Condition,y=flux,color = Condition, fill = Condition)) +
  geom_jitter() +
  geom_hline(yintercept = 0) +
  facet_wrap(~rxns_names,
             scales = "free") +
  #theme_ipsum_rc(axis = TRUE,
                 #axis_title_face = "bold") +
  ggtitle("Porphyrin and chlorophyll metabolism") +
  xlab("") +
  ylab("Flux (mmol/gDWh)") +
  scale_y_continuous(breaks = NULL) +
  theme_bw() +
  theme(legend.position = "bottom",
        aspect.ratio = 0.7,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(), 
        plot.title = element_text(face = "bold")) +
  scale_color_manual(values = wes_palette("Cavalcanti1", n = 3, type = "discrete")) +
  scale_fill_manual(values = wes_palette("Cavalcanti1", n = 3, type = "discrete")) +
  stat_compare_means(label = "p.signif",vjust = 2, hjust = 1)  -> p_porphyrin

  ggsave("p_porphyrin.png", device = "png", width = 30, height = 25, units = "cm", dpi = "retina")

  
  
```
 
```{r}
pathway_to_plot = "Pentose phosphate pathway"
 
df_long_nonzero %>%
  right_join(nonZero_fluxes, by = "rxns_names") %>%
  mutate(Condition = case_when(
    grepl("Glu",solution) ~ "Glu",
    grepl("PET",solution) ~ "PET",
    grepl("PU",solution) ~ "PU",
    )) %>%
  filter(pathway_id == pathway_to_plot) %>% 
  ggplot(aes(x=Condition,y=flux,color = Condition, fill = Condition)) +
  geom_jitter() +
  geom_hline(yintercept = 0) +
  facet_wrap(~rxns_names,
             scales = "free") +
  #theme_ipsum_rc(axis = TRUE,
                 #axis_title_face = "bold") +
  ggtitle("Pentose phosphate pathway") +
  xlab("") +
  ylab("Flux (mmol/gDWh)") +
  scale_y_continuous(breaks = NULL) +
  theme(legend.position = "none") +
  scale_color_manual(values = wes_palette("Cavalcanti1", n = 3, type = "discrete")) +
  scale_fill_manual(values = wes_palette("Cavalcanti1", n = 3, type = "discrete"))  -> p_ppp

  ggsave("p_ppp.png", device = "png", width = 30, height = 25, units = "cm", dpi = "retina")
```
 
 
Plot heatmap of DE fluxes
```{r}
# Extract nonzero fluxes
nonZero_fluxes = 
  df_long_nonzero %>% 
  group_by(rxns_names) %>% 
  summarize(zeroFlux = sum(flux) == 0) %>%
  ungroup() %>% 
  filter(zeroFlux == FALSE)
 
diffFluxes = 
  df_long_nonzero %>% 
  right_join(nonZero_fluxes, by = "rxns_names") %>%
  select(rxns_names, pathway_id) %>% 
  inner_join(res_wilcox, by = "rxns_names") %>% 
  distinct() %>%
  arrange(FDR) %>%
  filter(FDR < 0.05)
 
# Format for Heatmap
hm.data = df_long_nonzero %>% 
  right_join(nonZero_fluxes,by = "rxns_names") %>% 
  right_join(diffFluxes) %>% 
  select(rxns_names,solution,flux) %>% 
  pivot_wider(names_from = solution,
              values_from = flux) %>% 
  arrange(rxns_names) %>% 
  as.data.frame()
 
rownames(hm.data) = hm.data[,1]
hm.data = hm.data[,-1]
 
hm.data = as.matrix(hm.data)
 
# Get sample data for annotation

sample.cond =
  plotData %>% 
  as_tibble() %>% 
  mutate(Condition = case_when(
    grepl("Glu",solution) ~ "Glu",
    grepl("PET",solution) ~ "PET",
    grepl("PU",solution) ~ "PU",
    )) %>% 
    pull(Condition)
sample.cond = 
  df_long_nonzero_cond %>% 
  right_join(nonZero_fluxes,by = "rxns_names") %>% 
  right_join(diffFluxes) %>% 
  pull(Condition)
 
pathway.annot = df_long_nonzero %>% 
  right_join(nonZero_fluxes,by = "rxns_names") %>% 
  select(rxns_names,pathway_id) %>%
  distinct() %>% 
  arrange(rxns_names) %>% 
  pull(pathway_id)
 
# Create annotations for heatmap
ha.col = HeatmapAnnotation(Carbon_source = sample.cond,
                       col = list(Carbon_source = c(Glucose = "darkorange1", Polyurethane = "violetred", PET = "darkmagenta")))
 
 
# Draw heatmap
hm = Heatmap(hm.data,
        name = "scaled-centered intesities",
        clustering_distance_rows = "euclidean", 
        clustering_distance_columns = "euclidean",
        show_column_names = FALSE,
        show_row_names = FALSE,
        #top_annotation = ha.col,
        #left_annotation = ha.row,
        #col = col_fun,
        #column_dend_height = unit(25, "mm"),
        #row_dend_width = unit(25, "mm"),
        show_row_dend = TRUE,
        show_column_dend = TRUE)
 
hm
```
 
 Heatmap with tidyHeatmap
```{r}
df_long_nonzero_cond %>% 
  as_tibble() %>% 
  heatmap(rxns_names, solution, flux, .scale = "column") %>% add_tile(Condition)
```
 
 
```{r}
df_long_nonzero %>% 
  right_join(nonZero_fluxes,by = "rxns_names_fluxes") %>% 
  mutate(Condition = ifelse(grepl("Glu",solution),"Glucose","Polyurethane")) %>% 
  filter(rxns_names_fluxes == "ENO") %>% 
  group_by(Condition) %>% 
  summarise(median(flux))
 
df_long_nonzero %>% 
  right_join(nonZero_fluxes,by = "rxns_names_fluxes") %>% 
  mutate(Condition = ifelse(grepl("Glu",solution),"Glucose","Polyurethane")) %>% 
  filter(rxns_names_fluxes == "FESD1s") %>% 
  group_by(Condition) %>% 
  summarise(median(flux))
```




 