---
title: "VizFlux non optimal"
author: "iGEM Chalmers 2020"
output: html_notebook
---

load libraries 
```{r}
library(tidyverse)
library(cowplot)
library(ggpubr)
library(ggsci)
library(ggrepel)
library(ggExtra)
library(readxl)
library(ComplexHeatmap)
library(Rtsne)
library(janitor)
library(hrbrthemes)
library(wesanderson)
library(tidyHeatmap)
select = dplyr::select
mutate = dplyr::mutate
```


load data
```{r}
df_glu <- read_excel("/Users/castilln/Desktop/igem/github/GEM/scripts/scrap/randomSolutionsGlu_t.xlsx")
df_pet <- read_excel("/Users/castilln/Desktop/igem/github/GEM/scripts/scrap/randomSolutionsPET_t.xlsx")
df_pu <- read_excel("/Users/castilln/Desktop/igem/github/GEM/scripts/scrap/randomSolutionsPU_t.xlsx")
 
df_kegg = read_csv("/Users/castilln/Desktop/igem/github/GEM/Data/kegg_df.csv")
```

join flux distribution from different c sources
```{r}
#glu + pu
df_full = 
 full_join(df_glu,df_pu, by = "rxns_names")

#glu + pu + pet
df_full = full_join(df_full, df_pet, by="rxns_names")
head(df_full)
```

transform to long & parsing to make sure there are no 0 value
```{r}
  df_full %>% 
  pivot_longer(cols = -rxns_names,
               names_to = "solution",
               values_to = "flux") %>% 
  distinct() %>% 
  mutate(flux = ifelse(is.na(flux),0,flux)) -> df_long
```

select unique flux distributions 
```{r}
df_long %>% 
  group_by(rxns_names, flux) %>% 
  arrange(rxns_names, flux, solution) %>% 
  slice(1L) -> filter_sol

#select vector containing unique solutions names 
filter_sol %>% ungroup() %>%  select(solution) %>% distinct() -> filter_only_sol 

#merge with out dataframe and keep unique solutions 
df_long %>% inner_join(filter_only_sol, by = "solution") -> df_long_unique

```

keep solutions for which biomass is produced 
```{r}
nonZero_samples = 
  df_long_unique %>% 
  filter(rxns_names == "BIOMASS_Ec_iML1515_core_75p37M") %>% 
  filter(flux > 0)
 
head(nonZero_samples)
```

Filter out all non zero samples, add pathway information from KEGG
```{r}
df_long_nonzero = 
  nonZero_samples %>% 
  select(solution) %>% 
  left_join(df_long_unique, by = c("solution" = "solution")) %>% 
  left_join(df_kegg, by = c("rxns_names" = "rxn_id")) %>% 
  group_by(solution) %>% 
  mutate(NA_fluxes = ifelse(sum(is.na(flux)) > 0,TRUE,FALSE)) %>% 
  ungroup() %>% 
  filter(NA_fluxes == FALSE)
```

Convert into wide format
```{r}
df_wide_nonzero =
  df_long_nonzero %>% 
  select(-c(gene_id,pathway_id)) %>% 
  pivot_wider(names_from = solution,
              values_from = flux) %>% 
  as.data.frame() 
 
rownames(df_wide_nonzero) = df_wide_nonzero[,1]
df_wide_nonzero = df_wide_nonzero[,-1]
```

### Dimnesionality reduction
PCA
```{r}
pca1 = prcomp(t(df_wide_nonzero),center = TRUE)
 
plotData = pca1$x[,1:2]
plotData = cbind(solution = rownames(plotData),plotData)
rownames(plotData) = NULL
 
head(plotData)
 
sample_names = plotData[,1]
 
plotData %>% 
  as_tibble() %>% 
  mutate(Condition = case_when(
    grepl("Glu",solution) ~ "Glu",
    grepl("PU",solution) ~ "PU",
    grepl("PET",solution) ~ "PET" 
    )) %>% 
  filter(!is.na(Condition)) %>% 
  mutate(PC1 = as.double(PC1),
         PC2 = as.double(PC2))  -> plotData_gg
  ggplot(plotData_gg, aes(x = PC1,y = PC2, color = Condition)) +
  geom_point() 
  scale_y_continuous(breaks = NULL)   #to fix error
  
```

tSNE
```{r}
set.seed(6) #obtain always the same plot 
tsne1 = Rtsne(t(df_wide_nonzero),
             dims = 2,
             perplexity = 20,
             theta = 0.5,
             check_duplicates = FALSE)
 
tsne1$Y
 
plotData =
  tsne1$Y %>% 
  as_tibble() %>% 
  mutate(solution = sample_names)
 
plotData = 
  plotData %>% 
  mutate(Condition = case_when(
    grepl("Glu",solution) ~ "Glu",
    grepl("PET",solution) ~ "PET",
    grepl("PU",solution) ~ "PU",
    )) %>% 
  mutate(V1 = as.double(V1),
         V2 = as.double(V2))
 
plotData %>% 
  ggplot(aes(x = V1,y = V2, color = Condition, group = Condition)) +
  geom_point(size = 2) +
  xlab(element_blank()) +
  ylab(element_blank()) +
  theme_bw() +
  #ggtitle("tSNE") +
  stat_ellipse() +
  theme(legend.position = "bottom",
        aspect.ratio = 1,
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(face = "bold")) +
  scale_color_manual(values = wes_palette("Cavalcanti1", n = 3, type = "discrete")) -> tSNE_plot

  ggsave("tSNE_plot.png", device = "png", width = 15, height = 15, units = "cm", dpi = "retina")

```

Hierarchical clustering and heatmaps
```{r}
# Extract nonzero fluxes
nonZero_fluxes = 
  df_long_nonzero %>% 
  group_by(rxns_names) %>% 
  summarize(zeroFlux = sum(flux) == 0) %>%
  ungroup() %>% 
  filter(zeroFlux == FALSE)
 
# Format for Heatmap
hm.data = df_long_nonzero_cond %>% 
  right_join(nonZero_fluxes,by = "rxns_names") %>% 
  select(rxns_names,solution,flux) %>% 
  pivot_wider(names_from = solution,
              values_from = flux) %>% 
  arrange(rxns_names) %>% 
  as.data.frame()
 
rownames(hm.data) = hm.data[,1]
hm.data = hm.data[,-1]
 
hm.data = as.matrix(hm.data)
 
# Get sample data for annotation


sample.cond = 
  df_long_nonzero_cond %>% 
  right_join(nonZero_fluxes,by = "rxns_names") %>% 
  pull(condition)
 
pathway.annot = df_long_nonzero %>% 
  right_join(nonZero_fluxes,by = "rxns_names") %>% 
  select(rxns_names,pathway_id) %>%
  distinct() %>% 
  arrange(rxns_names) %>% 
  pull(pathway_id)
 
# Create annotations for heatmap

#ha.col = HeatmapAnnotation(Carbon_source = sample.cond,
                       col = list(Carbon_source = c("Glu" = "darkorange1", "PU" = "violetred", "PET" = "darkblue")))
 
# Draw heatmap
hm = Heatmap(hm.data,
        name = "scaled-centered intesities",
        clustering_distance_columns = "euclidean",
        clustering_distance_rows = "euclidean",
        show_column_names = FALSE,
        show_row_names = FALSE,
        top_annotation = ha.col,
        #left_annotation = ha.row,
        #col = col_fun,
        column_dend_height = unit(25, "mm"),
        row_dend_width = unit(25, "mm"),
        show_row_dend = F,
        show_column_dend = F)
 
hm
```
alternative: tidyHeatmap
Tidy heatmap
```{r}
df_long_nonzero_cond = 
  df_long_nonzero %>% 
  mutate(condition = case_when(
    grepl("Glu",solution) ~ "Glu",
    grepl("PET",solution) ~ "PET",
    grepl("PU",solution) ~ "PU",
    ))

df_random = sample_n(df_long_nonzero_cond, 100000)

Carbon_source = c("darkorange1","violetred", "darkblue")
names(Carbon_source) = c("Glu", "PU", "PET")


giveNAs = which(is.na(as.matrix(dist(df_random))),arr.ind=TRUE)

df_random %>% 
  group_by(condition) %>% 
  heatmap(
    rxns_names,
    solution,
    flux, 
    transform = NULL,
    .scale = "both",
    palette_value = circlize::colorRamp2(c(-20, -5, 0, 5, 20), viridis::magma(5)),
  ) %>% 
  add_tile(condition)
 

```

```{r}
df_long_nonzero_cond %>% 
  mutate(solution = gsub(" ","",solution)) %>%
 # rename("´solution´" = "solution") %>% 
  heatmap(.row = rxns_names, .column = solution, .value = flux,
          palette_value = circlize::colorRamp2(c(-2, -1, 0, 1, 2), viridis::magma(5))) %>% 
  add_tile(condition) %>% 
  add_tile(pathway_id)
```

Wilcoxon tests
```{r, warning=F}
options(scipen = 1)
res_wilcox =
  df_long_nonzero %>%
  right_join(nonZero_fluxes, by = "rxns_names") %>%
  mutate(condition = ifelse(grepl("Glu", solution), "Glucose", "Polyurethane")) %>%
  select(rxns_names, flux, condition) %>%
  group_by(condition, rxns_names) %>%
  summarise(flux = list(flux)) %>%
  pivot_wider(names_from = condition, values_from = flux) %>%
  group_by(rxns_names) %>%
  mutate(
    p_value = wilcox.test(unlist(Glucose), unlist(Polyurethane))$p.value,
    fold_change = sign(median(unlist(Polyurethane), median(unlist(Glucose)))) * 
      abs(median(unlist(Polyurethane)) - median(unlist(Glucose)))
  ) %>%
  select(rxns_names, p_value, fold_change) %>%
  mutate(FDR = p.adjust(p_value, method = "BH")) %>%
  filter(FDR < 0.05) %>%
  arrange(FDR)
```

Show wilcox test results
```{r}
df_long_nonzero %>% 
  right_join(nonZero_fluxes, by = "rxns_names") %>%
  select(rxns_names, pathway_id) %>% 
  inner_join(res_wilcox, by = "rxns_names") %>% 
  distinct() %>% 
  arrange(FDR)
```

Filter out those with pathways
```{r}
df_long_nonzero %>% 
  right_join(nonZero_fluxes, by = "rxns_names") %>%
  select(rxns_names, pathway_id) %>% 
  inner_join(res_wilcox, by = "rxns_names") %>% 
  distinct() %>%
  arrange(FDR) %>% 
  filter(!is.na(pathway_id)) %>% 
  arrange(fold_change)  #get the ones with largest fc first 
```


distribution of yield by condition
```{r}
df_long_nonzero_cond %>% 
  filter(rxns_names == "BIOMASS_Ec_iML1515_core_75p37M") %>%
  mutate(yield = case_when(
    grepl("Glu",solution) ~ (flux/1.8016),
    grepl("PET",solution) ~ flux/(6*0.1922),
    grepl("PU",solution) ~ flux/(3.5*0.3123)
    )) %>% 
  filter(flux < 60) -> yield_long_nonzero_cond
```


```{r}
yield_long_nonzero_cond %>%  
ggplot(aes(x=condition,y=flux,color = condition, fill = condition)) +
  geom_jitter() +
  #theme_ipsum_rc(axis = TRUE,
    #             axis_title_face = "bold") +
  xlab("") +
  ylab("Yield (g C source/gDWh)") +
  theme(legend.position = "none") +
  scale_color_manual(values = wes_palette("Cavalcanti1", n = 3, type = "discrete")) +
  scale_fill_manual(values = wes_palette("Cavalcanti1", n = 3, type = "discrete"))
```

